Davos: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2023-09-29 paper/main.tex

\textcolor{red}{\textbf{TODO: add note about default/fallback project for non-traditional notebook interfaces}}

>>>>

While Davos enables developers to conveniently specify all project
dependencies, there are some edge cases and limitations that are worth
considering. Many Python packages include (in their setup options) additional
dependencies that often carry their own version specifications. Although Davos
will check that the correct version of the requested top-level package is
installed and imported into the workspace, the version numbers of any
\textit{dependencies} of the requested package are not checked. In principle,
this could lead to unexpected behavior, for example if a given package's
dependencies (or dependencies of those dependencies, etc.) were left
under-specified. A developer could mitigate this by explicitly smuggling exact
version numbers of every project dependency (e.g., obtained via pip freeze).
However, for projects where the versions of dependencies of smuggled packages
also need to be precisely controlled, a lockfile or a \texttt{requirements.txt}
file produced by \texttt{pip freeze} (i.e., explicitly specifying \textit{all} packages'
version numbers) may provide a more comprehensive alternative to Davos.

Our Project infrastructure (Sec.~\ref{subsec:projects}) provides a ``safe'' way
of managing project dependencies without interfering with the user's Python
environment. However, our implementation of this functionality cannot
anticipate every possible development environment. As of this writing, we
support many common notebook environments, including classic Jupyter Notebooks,
Google Colaboratory, Kaggle Notebooks, Visual Studio Code, among others.
However, some environments (e.g., Visual Studio Code) implement their own
mechanisms for rendering and running notebook code. In cases where our Project
infrastructure is not equipped to handle a particular development or runtime
environment, Davos will fall back to installing packages in a ``default''
environment that is shared across all notebooks.  This enables Davos to protect against
modifying the runtime environment, but it also means that packages may be installed
or overwritten across notebooks that use different versions of those packages.
